---
title: "CloneVerify"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CloneVerify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is the CloneVerify package which uses 6 distinct SNPs to identify clones  (Criollo_22, Matina_1_6, Pound_7), when the tags fall so that we avoid collecting incorrect data. The package is dependent on Tidyverse; so it should also be installed. 
```{r}
#required package:
library(tidyverse)
library(CloneVerify)
```
Introduction:
The package requires a standard format for data. I have included an example of a correct data set called "demo". It is 68 rows and 10 columns wide. It is a group of trees of 3 possible clones (Criollo_22, Matina_1_6, Pound_7) that are most commonly used in clonal trials. The trees often lose labels or the tags find and often have to be re-identified. 

Each row in the data set represents an individual tree. The columns are: 

Chip.BC = The chip barcode the genotyping was done on. 

Well = The location of the sample on the chip

ID = The status of the tree. Whether it is a clonal reference (Criollo_22, Matina_1_6, Pound_7) or if it the sample/tree that is an unknown( we do not know what it is a clone of). Unknowns can happen for a lot of reasons such as loss of tag information, or poorly kept maps. I have created an R package that will help my analyze this data more efficiently for the future. 

DC= A unique identifier given to each sample/tree

Columns 5-10: Are 10 SNP marker names that are used for "offtyping" the trees. These specific markers are used for these three clones because they result in  1 = homozygote_1,  2= homozygote_2, 3= Heterozygote. So they make it really easy to tell these three clones apart. 
na = function(data) {
    na.omit(data)
}
```{r}
head(demo)
dim(demo)
```

Functions:
# na
As this is a real world data set I want to exclude missing data (NA). The first step of the method is to remove the rows with missing data. I did this with the na function in the package below:
na = function(data){
  na.omit(data)
}
```{r}
d2 = na(demo)
```
This removes all the samples/trees that have missing data on at least one of the SNPs. 

#uni_gt
This function will give the amount of unique genotypes:
uni_gt = function(data) {
  unique(data)
}
```{r}
uni_gt(d2[,c(5:10)])
```
There are 3 unique genotypes in the data set, meaning that all three clones are present in the sampling and no offtypes have occured.

#dend
the dend() fucntion will give a visual represenation to check that the data clusters into 3 groups 

dend = function (x ){
dend <- hclust(dist(x))
plot(dend)
}
```{r}
d2_sub <- d2[ , 5:10]
dend(x= d2_sub)
```
Again the dendogram show visually that there are 3 groups. 

#Verify
The Verify functions adds a column to the data (Verified_clone). Which will identify which clone each tree is.
Verify = function (data){
  Criollo_22 = data %>%
  filter(TcSNP606 == 1, ) %>%
  mutate(Verified_clone = "Criollo_22")
  
  Matina_1_6= data %>%
  filter(TcSNP606 == 2, ) %>%
  mutate(Verified_clone = "Matina_1_6")
  
  Pound_7 = data %>%
  filter(TcSNP606 == 3, ) %>%
  mutate(Verified_clone = "Pound_7")
  
  Verified_Clonal_Data = do.call("rbind", list(Criollo_22, Matina_1_6, Pound_7))
  
  head(print(Verified_Clonal_Data))
}

```{r}


Final_data = Verify(d2)

```


